---
title: "Available Datasets in R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                        fig.path = "images/")
```

For learning or testing purpose many a time we need some data in our R working environment. There are lots of built in data sets available with base R as well as the R packages we installed. We can view the list of data available in each packages as well as can make use of it.

> I am showing the packages available in my enviroment. You might or might not have these packages. Try to follow the steps with refer to your available packages.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
  
  #First check the list of data sets available in base R in the package datasets
  data()
  # Datasets available in all the installed packages
  data(package = .packages(all.available = TRUE))

  # Datasets in a particular package
  data(package = "caret")

  # Load data from base R package datasets
  data("mtcars")

  # Read description about the dataset
  ?mtcars

```

```{r}
# The data frame
  head(mtcars)

```

```{r, message=FALSE, warning=FALSE}
# OR
  df <- mtcars

  # OR df <- packagename::dataset_name
  df <- datasets::mtcars


  # Data from caret package
  data("dhfr",package = "caret")
```


Sometimes there might be difficulties while making a dataframe from some packages when the object is loaded as value.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
  data("gadmCHE",package = "leaflet")
  # Then create dataFrame using following
  df <- as.data.frame(gadmCHE)
```

