---
title: "Connecting MongoDB and querying data from R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path = "images/")
```

According to the documentation MongoDB is a document database with the scalability and flexibility that you want with the querying and indexing that you need. It is a NoSQL based leading database.  

> In this article I will be using MongoDB community edition installed in my 64 bit windows machine. If you want you can use your MongoDB database hosted in some server.   

If you are new to MongoDB and want to know more go through the following links :

* [Installation of community Edition](https://docs.mongodb.com/manual/installation/)
* [Official documentation](https://docs.mongodb.com/manual/)
* [Tutorials Point tutorial](https://www.tutorialspoint.com/mongodb/index.htm)  

> As of now you are done with your installation locally

There are many MongoDB client for R. I will be using **mongolite**, which is fast and simple.

**First of all, lets call the `mongolite` package from R :**
```{r, echo=TRUE, message=FALSE, warning=FALSE}
if(require(mongolite)){
    print("mongolite is loaded correctly")
} else {
    print("trying to install mongolite")
    install.packages("mongolite")
    if(require(mongolite)){
        print("mongolite installed and loaded")
    } else {
        stop("could not install mongolite")
    }
}
```

**Run local MongoDB :**

* Go-to ur MongoDB install path and open command prompt/or from command prompt go to your path: `eg: /usr/MongoDB/bin`
* Start MongoDB by `mongod` from your shell or command line
* Now your MongoDB is running locally in port `27017`

**Connect MongoDB local server :**

* Initiate a connection to MogoDB Server.
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# 1. Connection to Local Server initiated by mongod : 27017 is the port where mongodb server is running
connectionLocal <- mongo(url = "mongodb://localhost:27017")
print(connectionLocal)
```

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# 2. Connecting to a remote server running in some host and port
# mongodb://[username:password@]host1[:port1][,host2[:port2],...[/[database][?options]]

# Let us consider we have a collection 'income-data' in 'people-database' running in a server (192.0.something.something or some-domain)

# connection <- mongo(url = "mongodb://my-username:my-password@somedomain:port")
# commented it, because I don't really have such user name and password

# Rather conncting to a public data-base
connection <- mongo("flights", url = "mongodb://readwrite:test@ds043942.mongolab.com:43942/jeroen_test")
# -> here collection is 'flights' and database 'jeroen_test'; username:readwrite; password:test; host?:ds043942.mongolab.com; port:jeroen_test
```

* Create a collection and Insert Data
```{r}
# Get some data from "nycflights" R package: Let "flights" data
data("flights", package = "nycflights13")
head(flights, 5)
# By default if we don't mention the collection & db while connecting MongoDB server it is taken as 'test' and 'test' : Lets define that in connectionLocal object 
connectionLocal <- mongo(collection = "fligts-data",db = "Rmongolite",url = "mongodb://localhost:27017")
# Insert AirQuality data
connectionLocal$insert(flights)
# If you want to put same data again in the same collection (just to check): you can just write the above command again

# We successfully inserted the data
# Now you can view the data in the DB and Collection in MongoDB management tolls like 'Robomongo'

# Check How many rows are available to varify (I am running the same code again and again; so the count may be original(ie.336776)*somenumber(1 or 2 or more))
connectionLocal$count()
```
* Query data from MongoDB : you can query all the available data in a collection or query specific subset of data from the collection. (MongoDB uses **JSON based Syntax** to query data)
  
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# 1. Get all the from a collection: our flights collection is connected by ConnectionLocal object
#connectionLocal$find()
# OR : empty '{}' means get all the data
#connectionLocal$find('{}')

# 2. Read all the Data into R (create a R DataFrame)
df <- connectionLocal$find()
head(df,5)

# 3. Query by using some logical expression. Get the data where : carrier=="UA" and distnace>1000
dfq1 <- connectionLocal$find('{"carrier":"UA","distance" : {"$gt":1000}}')
head(dfq1,5)

# 4. Filter Fields : Get only few required fields (let: arr_delay,carrier,origin,dest,distance) combine with the previous query
dfq2 <- connectionLocal$find(
  query = '{"carrier":"UA","distance" : {"$gt":1000}}',
  fields = '{"arr_delay":true, "carrier":true, "origin":true, "dest" : true, "distance":true}'
  )
head(dfq2,5)
# while filtering random UUID will be created automatically, If you don't want it need to explicitly mention -> "_id" : false

# 5. Sort and limit : limit is for restrict the number of outputs(let 10), lets sort in desscending order
dfq3 <- connectionLocal$find(
  query = '{"carrier":"UA","distance" : {"$gt":1000}}',
  fields = '{"_id":false,"arr_delay":true, "carrier":true, "origin":true, "dest" : true, "distance":true}',
  sort = '{"distance" : -1}',
  limit = 15
  )
# Check the dimension of the DataFrame
dim(dfq3)
head(dfq3,5)

# 6. Select by Date : greater than 2013-08-25T18:00:00Z
dfq4 <- connectionLocal$find(
  query = '{"time_hour": { "$gte" : { "$date" : "2013-08-25T18:00:00Z" }}}',
  fields = '{"time_hour" : true, "carrier" : true, "origin":true, "_id": false}'
)
head(dfq4,5)

# Many other operations like indexing, iterating are also can be performed -> explore yourself :)
```

* Aggregation of data using aggregate() method
```{r, echo=TRUE, message=FALSE, warning=FALSE}
dfagr1 <- connectionLocal$aggregate(
  '[{"$group":{"_id":"$carrier", "count": {"$sum":1}, "average":{"$avg":"$distance"}}}]',
  options = '{"allowDiskUse":true}'
)

head(dfagr1,5)
```

* MapReduce : map and reduce operations by giving custom functions in JavaScript
```{r}
dfmr1 <- connectionLocal$mapreduce(
  map = "function(){emit(Math.floor(this.distance/100)*100, 1)}", 
  reduce = "function(id, count){return Array.sum(count)}"
)
colnames(dfmr1) <- c("scale_distance", "counts")
head(dfmr1,5)
```

*There are many more functionalities we can perform in MongoDB from R.
> Drop database/collection : connectionLocal$drop()
> Remove some rows : connectionLocal$remove('{"Carrier":"UA"}')
> Remove only one row : connectionLocal$remove('{"Carrier":"UA"}', just_one = TRUE)
> Remove all records but not the collection : connectionLocal$remove()
> We can also import, expoer **JSON, BSON** files
> Update a single row/document : connectionLocal$update('{"carrier":"UA"}', '{"$set":{"distance": 1111}}')
> Update entire rows/documents : connectionLocal$update('{}','{"carrier":"UA"}', '{"$set":{"distance": 1111}}',multiple = TRUE)
> Add if no matches found : connectionLocal$update('{"carrier":"UA"}', '{"$set":{"distance": 1111}}', upsert = TRUE)
